---
title: "Fitness by GDP/Cap"
author: "SBP"
date: "2/29/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(here)
library(tidyverse)
library(countrycode)
library(haven)
library(pwt9)
library(WDI)
library(vroom)
```


- [ ] Klargør GDP data
- [ ] Klargør population data
- [ ] Klargør pop/cap data
- [ ] Klargør Trade data
- [ ] Klargør fitness data

## Reference year

```{r}
ref_year <- 2013
```

## Prepare Fitness data

```{r}
fitness_df <- vroom("/home/post/university/humgeo_thesis/data/downloaded/Economic_Fitness_CSV/Data.csv")

fitness_tidy <- fitness_df %>%
  rename(country_code = `Country Code`, metric = `Indicator Name`) %>%
  select(-c(`Country Name`, `Indicator Code`)) %>%
  gather(-c(country_code, metric), key = year, value = fitness) %>%
  mutate(year = as.numeric(year)) %>%
  filter(metric == "Economic Fitness Metric" & year == ref_year) %>%
  select(-metric)
```

## Prepare Economic complexity data

```{r}
econ_comp_df <- vroom("/home/post/university/humgeo_thesis/data/downloaded/Economic Complexity/eci_country_rankings_2013_2017.csv") %>%
  mutate(country_code = countrycode(Country, 'country.name', 'iso3c')) %>%
  select(country_code, year = Year, ECI, ECI_plus = `ECI+`) %>%
  filter(year == ref_year)
```


## Create list of countries to sort GDP and trade data by

```{r}
fit_country_set <- fitness_tidy %>%
  distinct(country_code) %>%
  pull()

eco_country_set <- econ_comp_df %>%
  distinct(country_code) %>%
  pull()

country_set <- intersect(eco_country_set, fit_country_set)
```

## Prepare GDP data
The WDI package provides the WDI function, which uses an API to source the data from the indicators. The GDP data is measured in PPP (constant 2011 international $). Link: https://data.worldbank.org/indicator/NY.GDP.MKTP.PP.KD.

```{r}
wdi_gdp <- as_tibble(
     WDI(
       country = "all",
       indicator = "NY.GDP.MKTP.PP.KD",
       extra = TRUE,
       start = 2010)
   ) 

wdi_gdp_tidy <- wdi_gdp %>% 
  select(
    country_code = iso3c, year, gdp = NY.GDP.MKTP.PP.KD
  ) %>%
  filter(country_code %in% country_set & year == ref_year) %>%
  mutate(country_code = as.character(country_code))

```

The variable NY.GDP.PCAP.PP.CD is in GDP per capita, PPP (current international $). This variable is national accounts-based (https://datatopics.worldbank.org/world-development-indicators/themes/economy.html).

```{r}
wdi_gdp_cap <- as_tibble(
     WDI(
       country = "all",
       indicator = "NY.GDP.PCAP.PP.CD",
       extra = TRUE,
       start = 2010)
   )

wdi_gdp_cap_tidy <- wdi_gdp_cap %>% 
  select(
    country_code = iso3c, year, gdp_cap = NY.GDP.PCAP.PP.CD
  ) %>%
  filter(country_code %in% country_set & year == ref_year) %>%
  mutate(country_code = as.character(country_code))

gdp_cap <- read_csv("/home/post/university/SGED07/assignments/papers/final_folder/data/API_NY.GDP.PCAP.PP.KD_DS2_en_csv_v2_867594/gdp_cap_ppp_2011_int_doll.csv",
                       skip = 4) %>%
  select(-c(`Country Name`, `Indicator Name`, `Indicator Code`, X65)) %>%
  rename(iso3c = `Country Code`) %>%
  gather(-iso3c, key = year, value = gdp_cap) %>%
  mutate(year = as.numeric(year))
```

## Prepare population data

## Prepare trade data

```{r}
sitc_df <- vroom("/home/post/university/humgeo_thesis/data/downloaded/trade/country_sitcproduct2digit_year.tab", delim = "\t")

sitc_tidy <- sitc_df %>%
  select(
    country_code = location_code, year, sitc_code = sitc_product_code, export_value
    ) %>%
  filter(year == ref_year)
```

## Create list raw natural resource code list

1. Fish (SITC division 03)
2. Raw materials (SITC divisions 21, 23, 24, 25, 26, 29)
of which: Raw hides, skins and furskins (21), Crude rubber (23), Cork and wood (24), Wood pulp (25), Textile fibres (26), Crude animal and vegetable materials, not elsewhere specified (29)
3. Fuels and mining products (SITC section 3 and divisions 27, 28, 68) of which: Ores and other minerals (SITC divisions 27, 28), Fuels (SITC section 3), Non-ferrous metals (SITC division 68). 

"The sum of the above three product groups provides us with a basic, usable statistical definition of natural resources." 

```{r}
natural_res_set <- c("03", "21", "23", "24", "25", "26", "29", "27", "28", "68", "32", "33", "34", "35")

natural_res_export <- sitc_tidy %>% 
  filter(sitc_code %in% natural_res_set) %>%
  group_by(country_code, year) %>%
  summarize(
    total_nat_res_export = sum(export_value)
  ) %>%
  arrange(total_nat_res_export)

```

## Find natural resource export % of GDP

```{r}
natural_res_countries <- wdi_gdp_tidy %>%
  left_join(natural_res_export) %>%
  mutate(nat_res_of_gdp_perc = total_nat_res_export / gdp * 100) %>%
  mutate(nat_res_dummy = ifelse(nat_res_of_gdp_perc >= 10, 1, 0))

```

## Join datasets

```{r}
joined_tidy <- natural_res_countries %>%
  left_join(fitness_tidy) %>%
  left_join(econ_comp_df) %>%
  left_join(wdi_gdp_cap_tidy) %>%
  filter(!is.na(gdp)) 
```

## Create figure

```{r}
joined_tidy

lm(gdp_cap ~ fitness + nat_res_dummy, data = joined_tidy) %>%
  summary()

lm(log(gdp_cap) ~ ECI + nat_res_dummy, data = joined_tidy) %>%
  summary()

lm(log(gdp_cap) ~ ECI_plus + nat_res_dummy, data = joined_tidy) %>%
  summary()

lm(log(gdp_cap) ~ ECI, data = joined_tidy %>% filter(nat_res_dummy != 1)) %>%
  summary()

lm(log(gdp_cap) ~ ECI_plus, data = joined_tidy %>% filter(nat_res_dummy != 1)) %>%
  summary()


ggplot(joined_tidy, aes(y = log(gdp_cap), x = ECI, label = country_code, col = as.factor(nat_res_dummy))) +
  geom_text() +
  geom_smooth(method = "lm", se = FALSE)
  

ggplot(joined_tidy, aes(y = log(gdp_cap), x = ECI_plus, label = country_code, col = as.factor(nat_res_dummy))) +
  geom_text() +
  geom_smooth(method = "lm", se = FALSE)

```

```{r}

```


