# International trade
REMOVE SERVICE PRODUCTS

The data on international trade is compiled by UN COMTRADE and distributed by the Growth Lab at Harvard University. I run the analysis with two different classifications of products, the Harmonized System revision 92 (HS92) citep:the_growth_lab_at_harvard_university_international_2019 and the Standard International Trade Classification revision (SITC rev. 2) cite:the_growth_lab_at_harvard_university_international_2019-1.

## Cleaning procedure: SITC2 
### Download file

The file is provided by the Harvard Dataverse and is available for download here: https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/H8SFD2 (as "country_sitcproduct4digit_year.tab").

For consistency across all functions and data sets, I rename the year-variable to ~time~, the country-variable to ~region~, the product variable to ~unit~, and the export value-variable to ~intensity~.

```{R} 
library(here)
library(tidyverse)
library(vroom)

sitc_path <- here("../data/downloaded/trade/country_sitcproduct4digit_year.tab")
sitc_raw <- vroom(sitc_path, delim = "\t")

sitc <- sitc_raw %>%
	select(
        time = year,
        region = location_code,
        unit = sitc_product_code,
        intensity = export_value
    )
```
### Time independent filters

Following the Atlas of Economic complexity, four time-independent filters are used to filter the SITC-data.

The filters are:
 - countries must have data in 2010
 - countries must have a total export value > 1 bill in 2010
 - countries must have a population of at least 1.25 mio in 2010
 - countries must not be Tchad (TCD), Iraq (IRQ), Afghanistan (AFG)
 - products must not be service products: I remove products: "ZZZZ", "travel", "transport", "ict", "financial".

```{R}
ref_year <- 2010
ref_export <- 1000000000
ref_pop <- 1250000
ref_reliable <- c("AFG", "TCD", "IRQ")
unwanted_prods <- c("ZZZZ", "travel", "transport", "ict", "financial")
```


### Availability filter

I first create a set of countries that have data available in the reference year. This removes no-longer present countries, which helps for continuity in the time-series, but is a bit of an issue in earlier periods, given the prominence of fx the Soviet Union.

```{R}
available_set <- filter(sitc, time == ref_year) %>%
    pull(region) %>%
    unique()
```

### Export filter

To create the set of countries that clear the export filter, I use the trade data itself.

```{R}
export_set <- sitc %>%
	filter(time == ref_year) %>%
	group_by(region) %>%
	summarise(
        total_export = sum(intensity)
    ) %>%
	filter(total_export >= ref_export) %>%
	pull(region) %>%
    unique()
```

### Population filter

I use the World Development Indicators (WDI) population data to filter for minimum population in the reference year. The WDI distributes a package of the same name, from which I download the data.

```{R}

library(WDI)

pop_raw <- WDI(country = "all", indicator = "SP.POP.TOTL", extra = TRUE) %>%
	as_tibble()

pop_data <- pop_raw %>%
	mutate_if(is.factor, as.character) %>%
	mutate(
        region = str_to_upper(iso3c)
    ) %>%
	select(
        time = year,
        region,
        pop = SP.POP.TOTL
    )

pop_set <- pop_data %>%
	filter(time == ref_year & pop >= ref_pop) %>%
	pull(region) %>%
    unique()
```

### Reliability filter
The Atlas lists a few countries with unreliable trade data. I exclude these
countries.

```{R}
reliable_set <- sitc %>%
	filter(!(region %in% ref_reliable)) %>%
	pull(region) %>%
	unique()
```

### DONE Product filter

```{R}
product_set <- sitc %>%
  filter(!(unit %in% unwanted_prods)) %>%
  distinct(unit)
```
### Apply filters

I now apply the filters from above.
```{R}
sitc_filtered <- sitc %>%
	filter(region %in% export_set) %>%
	filter(region %in% pop_set) %>%
	filter(region %in% available_set) %>%
	filter(region %in% reliable_set) %>%
  filter(unit %in% product_set)
```

### Write file

Finally I write the file filtered only by the time-independent filtered into the data/prepared/ folder as hs92_filtered_ti.csv.

```{R}
write_csv(
  sitc_filtered,
  here("../data/prepared/sitc_filtered_ti.csv")
)
```


### TODO Time dependent filters
- [ ] Still need to check that all of the code works as advertised.

  Albaeik et al ("Improving the Economic Complexity Index", 2017) uses three time-dependent filters, in addition to the time-independent filters above. These three filters are:

 - exclusion of products for which, in a given year, 80% of countries have an export value of 0.
 - exclusion of countries that, in a given year, have an export value of 0 for 95% of products
 - exclusion of products with a global export value of less than 10 mio in a given year
 - lastly, all country-product export values of less than 5000 are set to 0.

```{r}
product_prop_limit <- 0.80
region_prop_limit <- 0.95
product_export_min <- 10000000
unwanted_prods <- c("ZZZZ", "travel", "transport", "ict", "financial")
```

### TODO First time-dependent filter

```{r}
product_summary <- sitc_filtered %>%
    group_by(time, unit) %>%
    summarise(
        regions_with_0_export = sum(intensity == 0),
        total_regions = n(),
        proportion_0 = regions_with_0_export / total_regions
    )

unit_year_exclusions <- product_summary %>%
    filter(proportion_0 >= product_prop_limit)
```

### TODO Second time-dependent filter

```{r}
region_summary <- sitc_filtered %>%
    group_by(time, region) %>%
    summarise(
        units_with_0_export = sum(intensity == 0),
        total_units = n(),
        proportion_0 = units_with_0_export / total_units
    )

region_year_exclusions <- region_summary %>%
    filter(proportion_0 >= region_prop_limit)
```

### TODO Third time-dependent filter

```{r}
global_product_export_vals <- sitc_filtered %>%
    group_by(time, unit) %>%
    summarise(
        total_export_val = sum(intensity)
    )

global_product_exclusions <- global_product_export_vals %>%
    filter(total_export_val < product_export_min)
```

### TODO Apply filters

I remove excluded observations (i.e. region-unit pairs) by using anti-join. anti-join(x, y) keeps all columns from x, and all the observations from x that does not have a match in y.

```{r}
sitc_filtered_td <- anti_join(sitc_filtered, unit_year_exclusions) %>%
    anti_join(region_year_exclusions) %>%
    anti_join(global_product_exclusions)
```

### TODO Write file

```{r}
output_path <- here("../data/prepared/sitc2_4digit_both_filters.csv")
write_csv(sitc_filtered_td, output_path)
```

## TODO Cleaning procedure: HS92 [14/14]

### Download file

The file is provided by the Harvard Dataverse and is available for download here: https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/T4CHWJ (as "country_hsproduct4digit_year.tab").

For consistency across all functions and data sets, I rename the year-variable to ~time~, the country-variable to ~region~, the product variable to ~unit~, and the export value-variable to ~intensity~.

```{R}
rm(list = ls())
library(here)
library(tidyverse)
library(vroom)

hs_path <- here("../data/downloaded/trade/country_hsproduct4digit_year.tab")
hs_raw <- vroom(hs_path, delim = "\t")

hs <- hs_raw %>%
	select(
        time = year,
        region = location_code,
        unit = hs_product_code,
        intensity = export_value
    )
```

### DONE Time independent filters

Following the Atlas of Economic complexity, I use four time-independent filters to filter the HS92-data.

The filters are:
 - countries must have data in 2010
 - countries must have a total export value > 1 bill in 2010
 - countries must have a population of at least 1.25 mio in 2010
 - countries must not be Tchad (TCD), Iraq (IRQ), Afghanistan (AFG)
 - products must not be service products: I remove products: "9999" (unknown), "XXXX", "travel", "transport", "ict", "financial".

```{R}
ref_year <- 2010
ref_export <- 1000000000
ref_pop <- 1250000
ref_reliable <- c("AFG", "TCD", "IRQ")
unwanted_prods <- c("9999", "XXXX", "travel", "transport", "ict", "financial")
```


### DONE Availability filter

I first create a set of countries that have data available in the reference year. This removes no-longer present countries, which helps for continuity in the time-series, but is a bit of an issue in earlier periods, given the prominence of fx the Soviet Union.

```{R}
available_set <- filter(hs, time == ref_year) %>%
    pull(region) %>%
    unique()
```

### DONE Export filter

To create the set of countries that clear the export filter, I use the trade data itself.

```{R}
export_set <- hs %>%
	filter(time == ref_year) %>%
	group_by(region) %>%
	summarise(
        total_export = sum(intensity)
    ) %>%
	filter(total_export >= ref_export) %>%
	pull(region) %>%
    unique()
```

### DONE Population filter

I use the World Development Indicators (WDI) population data to filter for minimum population in the reference year. The WDI distributes a package of the same name (that uses an API), from which I download the data.

```{R}
library(WDI)

pop_raw <- WDI(country = "all", indicator = "SP.POP.TOTL", extra = TRUE) %>%
	as_tibble()

pop_data <- pop_raw %>%
	mutate_if(is.factor, as.character) %>%
	mutate(
        region = str_to_upper(iso3c)
    ) %>%
	select(
        time = year,
        region,
        pop = SP.POP.TOTL
    )

pop_set <- pop_data %>%
	filter(time == ref_year & pop >= ref_pop) %>%
	pull(region) %>%
    unique()
```

### DONE Reliability filter
The Atlas lists a few countries with unreliable trade data. I exclude these countries.

```{R}
reliable_set <- hs %>%
	filter(!(region %in% ref_reliable)) %>%
	pull(region) %>%
	unique()
```

### DONE Product filter

```{R}
product_set <- hs %>%
  filter(!(unit %in% unwanted_prods)) %>%
  distinct(unit)
```

### DONE Apply filters

I now apply the filters from above.

```{R}
hs_filtered <- hs %>%
	filter(region %in% export_set) %>%
	filter(region %in% pop_set) %>%
	filter(region %in% available_set) %>%
	filter(region %in% reliable_set) %>%
  filter(unit %in% product_set$unit)
```

### Write file

Finally I write the file filtered only by the time-independent filtered into the data/prepared/ folder as hs92_filtered_ti.csv.

```{R}
write_csv(
  hs_filtered,
  here("../data/prepared/hs92_filtered_ti.csv")
)
```

### DONE Time dependent filters
- [ ] Still need to check that all of the code works as advertised.

Albaeik et al ("Improving the Economic Complexity Index", 2017) uses three time-dependent filters, in addition to the time-independent filters above. These three filters are:

 - exclusion of products for which, in a given year, 80% of countries have an export value of 0.
 - exclusion of countries that, in a given year, have an export value of 0 for 95% of products
 - exclusion of products with a global export value of less than 10 mio in a given year
 - lastly, all country-product export values of less than 5000 are set to 0.

```{R}
product_prop_limit <- 0.80
region_prop_limit <- 0.95
product_export_min <- 10000000
```

### DONE First time-dependent filter

```{R}
product_summary <- hs_filtered %>%
    group_by(time, unit) %>%
    summarise(
        regions_with_0_export = sum(intensity == 0),
        total_regions = n(),
        proportion_0 = regions_with_0_export / total_regions
    )

unit_year_exclusions <- product_summary %>%
    filter(proportion_0 >= product_prop_limit)
```

### DONE Second time-dependent filter

```{R}
region_summary <- hs_filtered %>%
    group_by(time, region) %>%
    summarise(
        units_with_0_export = sum(intensity == 0),
        total_units = n(),
        proportion_0 = units_with_0_export / total_units
    )

region_year_exclusions <- region_summary %>%
    filter(proportion_0 >= region_prop_limit)
```

### DONE Third time-dependent filter


```{R}
global_product_export_vals <- hs_filtered %>%
    group_by(time, unit) %>%
    summarise(
        total_export_val = sum(intensity)
    )

global_product_exclusions <- global_product_export_vals %>%
    filter(total_export_val < product_export_min)
```

### DONE Apply filters

I remove excluded observations (i.e. region-unit pairs) by using anti-join. anti-join(x, y) keeps all columns from x, and all the observations from x that does not have a match in y.

```{R}
hs_filtered_time_dep <- anti_join(hs_filtered, unit_year_exclusions) %>%
    anti_join(region_year_exclusions) %>%
    anti_join(global_product_exclusions)
```

### DONE Write file

I save the final filtered file in data/prepared/ as hs92_filtered_ti_and_td.csv.

```{R}
output_path <- here("../data/prepared/hs92_filtered_ti_and_td")
write_csv(hs_filtered_time_dep, output_path)
```

